SafeNet: A European database of plant-pollinator interactions

### 2024-03-21 Manuscript/Ms.qmd
Edits on Figure 1

@@

#Try to plot both together (same plot)
map = ggplot(countries) +
geom_sf(fill="floralwhite",color = "black", 
        alpha = 1,linewidth=0.1) + 
geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
           size=Counts, starshape = Group, color= Group))+  
coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
         xlim = c(-14, 39), ylim = c(37, 70))+
theme(panel.grid.major = element_blank(), 
      legend.position = ("top"),
      legend.box = "vertical",
      legend.background = element_blank(),
      legend.box.background = element_blank(),
      legend.key = element_blank(), 
      legend.title = element_text(face ="bold",  size = 6),
      legend.box.spacing = unit(0.2, "cm"),
      legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,-10,-10),
      axis.text = element_text(size=6,                  face="bold"),
      axis.title = element_text(face="bold", size = 8),
         legend.text = element_text(size = 6),                           legend.direction="horizontal",
         legend.spacing.y = unit(-0.3, "cm"),
      legend.title.align=0.5) +
theme(panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(colour = "black",
        fill=NA, size=1)) + 
 scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
scale_color_manual(values=c("#009E73", "#D55E00"),
         guide="none") +
scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
             guide=guide_legend(reverse=TRUE,
             override.aes=list(
                 size=2,
                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
  ylab("Latitude") +
xlab("Longitude") +
scale_x_continuous(breaks = seq(0, 20, by = 10)) +
scale_y_continuous(breaks = seq(40, 60, by = 10)) +
guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
ggtitle("(a)") +
theme(plot.title = element_text(vjust = -12, hjust=0.2, size=10))

@@

#Plot
p1 = ggplot(poll_spread, 
aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
geom_bar(stat = "identity", 
           color = "#D55E00",
           lwd = 0.015, fill = "#D55E00") +
ylab("Shared pollinators (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
  axis.ticks.x=element_blank(),
  axis.ticks.y = element_line(size = 0.2),
  plot.title = element_text(size = 8, face = "bold"), 
  axis.text.y = element_text(size=5),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
xlab(NULL) +
ylab(NULL) + 
ggtitle("Pollinators")+
geom_vline(xintercept = quantile_50, linetype="dashed", 
color = "black", size=0.25) + 
annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")

#----------------#
#Plants
#----------------#
plant_spread = data %>% 
select(Plant_rank, Plant_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
filter(Plant_rank == "SPECIES") %>% 
distinct(Plant_accepted_name) %>% 
ungroup() %>% 
count(Plant_accepted_name) %>% 
rename(n_studies_shared = n) %>% 
arrange(-n_studies_shared) %>% 
mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))

#Seelect quantile 50% from the total number of species
spp_number = plant_spread %>%  
select(Plant_accepted_name) %>% 
n_distinct()
df = seq(1, spp_number)
quantile_50 = quantile(df, 0.5)

#Obtain most common plant for plotting
most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
select(Plant_accepted_name) %>% pull()

#Plot
p2 = ggplot(plant_spread, 
aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
geom_bar(stat = "identity", 
           color = "#009E73",
           lwd = 0.015, fill = "#009E73") +
theme_minimal()+
theme(axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         axis.ticks.y = element_line(size = 0.2),
         axis.title.x = element_text(face = "bold", size = 7),            plot.title = element_text(size = 8, face = "bold"),
         axis.text.y = element_text(size=5),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_line(size=0.2, linetype =  2),
      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
xlab("Species") +
ylab("Occurrence across studies") + 
ggtitle("Plants") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
          color = "black", size=0.25) + 
annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
      ,color="black",  fontface = 'italic', size = 1.5)+
annotate(geom="text", x=20, y=-0.08, label="Common spp."
      ,color="black",  size = 1.5) +
annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
      ,color="black",  size = 1.5) + 
scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, max(plant_spread$Percent_total)), clip="off")

@@

p3 = ggplot(data2, aes(Year, Counts)) +
geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
theme_light() +
theme(plot.margin=unit(c(0,0,0,0), "mm"),
  axis.title.y = element_text(face = "bold", size = 7, vjust = -1.1),
  axis.title.x = element_text(face = "bold", size = 8),
  axis.text.y = element_text(size = 6),
  axis.text.x = element_text(size = 6,  angle = 90), 
  strip.text.x = element_text(size=12),
   axis.ticks.x = element_line(color="black"),
  strip.background = element_rect(
  color="black", fill=NA, size=0.5, linetype="solid")) + 
ylab("Number of studies") +
coord_cartesian(expand = F, clip = "off") + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
panel.border = element_blank(),
axis.line = element_line(color = 'black')) +
scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
ggtitle("(b)") +
theme(plot.title = element_text(size = 10,hjust = -0.35, vjust = 0.8))

plot2 = p3 / p1 / p2

#Plot everything
map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.58, 0.4))

>>>>

#Fix dates
data = data %>%
dplyr::mutate(Year = lubridate::year(Date), 
                Month = lubridate::month(Date), 
                Day = lubridate::day(Date))

@@

#Try to plot both together (same plot)
map = ggplot(countries) +
geom_sf(fill="floralwhite",color = "black", 
        alpha = 1,linewidth=0.1) + 
geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
           size=Counts, starshape = Group, color= Group))+  
coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
         xlim = c(-14, 39), ylim = c(37, 70))+
theme(panel.grid.major = element_blank(), 
      legend.position = ("top"),
      legend.box = "vertical",
      legend.background = element_blank(),
      legend.box.background = element_blank(),
      legend.key = element_blank(), 
      legend.title = element_text(face ="bold",  size = 6),
      legend.box.spacing = unit(0.2, "cm"),
      legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,-10,-10),
      axis.text = element_text(size=6,                  face="bold"),
      axis.title = element_text(face="bold", size = 8),
         legend.text = element_text(size = 6),                           legend.direction="horizontal",
         legend.spacing.y = unit(-0.3, "cm"),
      legend.title.align=0.5) +
theme(panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(colour = "black",
        fill=NA, size=1)) + 
 scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
scale_color_manual(values=c("#009E73", "#D55E00"),
         guide="none") +
scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
             guide=guide_legend(reverse=TRUE,
             override.aes=list(
                 size=2,
                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
  ylab("Latitude") +
xlab("Longitude") +
scale_x_continuous(breaks = seq(0, 20, by = 10)) +
scale_y_continuous(breaks = seq(40, 60, by = 10)) +
guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
ggtitle("(a)") +
theme(plot.title = element_text(vjust = -5, hjust=0.057, size=10))

@@

#Plot
#p1 = ggplot(poll_spread, 
#aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
#geom_bar(stat = "identity", 
#           color = "#D55E00",
#           lwd = 0.015, fill = "#D55E00") +
#ylab("Shared pollinators (%)") +
#theme_minimal()+
#theme(axis.text.x = element_blank(),
#  axis.ticks.x=element_blank(),
#  axis.ticks.y = element_line(size = 0.2),
#  plot.title = element_text(size = 8, face = "bold"), 
#  axis.text.y = element_text(size=5),
#  panel.grid.major.x = element_blank(),
#  panel.grid.minor.y = element_blank(),
#  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
#xlab(NULL) +
#ylab(NULL) + 
#ggtitle("Pollinators")+
#geom_vline(xintercept = quantile_50, linetype="dashed", 
#color = "black", size=0.25) + 
#annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = #1.5) +
#annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  #fontface = 'italic', size = 1.5)+
#scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
#annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
#coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, #max(poll_spread$Percent_total)), clip="off")
#
##----------------#
##Plants
##----------------#
#plant_spread = data %>% 
#select(Plant_rank, Plant_accepted_name, Study_id) %>% 
#group_by(Study_id) %>% 
#filter(Plant_rank == "SPECIES") %>% 
#distinct(Plant_accepted_name) %>% 
#ungroup() %>% 
#count(Plant_accepted_name) %>% 
#rename(n_studies_shared = n) %>% 
#arrange(-n_studies_shared) %>% 
#mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
#  
##Seelect quantile 50% from the total number of species
#spp_number = plant_spread %>%  
#select(Plant_accepted_name) %>% 
#n_distinct()
#df = seq(1, spp_number)
#quantile_50 = quantile(df, 0.5)
#
##Obtain most common plant for plotting
#most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
#select(Plant_accepted_name) %>% pull() 
#
#Plot
#p2 = ggplot(plant_spread, 
#aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
#geom_bar(stat = "identity", 
#           color = "#009E73",
#           lwd = 0.015, fill = "#009E73") +
#theme_minimal()+
#theme(axis.text.x = element_blank(),
#         axis.ticks.x=element_blank(),
#         axis.ticks.y = element_line(size = 0.2),
#         axis.title.x = element_text(face = "bold", size = 7),            plot.title = #element_text(size = 8, face = "bold"),
#         axis.text.y = element_text(size=5),
#         panel.grid.major.x = element_blank(),
#         panel.grid.minor.y = element_blank(),
#         panel.grid.major.y = element_line(size=0.2, linetype =  2),
#      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
#xlab("Species") +
#ylab("Occurrence across studies") + 
#ggtitle("Plants") + 
#geom_vline(xintercept = quantile_50, linetype="dashed", 
#          color = "black", size=0.25) + 
#annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
#annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
#      ,color="black",  fontface = 'italic', size = 1.5)+
#annotate(geom="text", x=20, y=-0.08, label="Common spp."
#      ,color="black",  size = 1.5) +
#annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
#      ,color="black",  size = 1.5) + 
#scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = #label_number(accuracy = 0.01)) +
#coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, #max(plant_spread$Percent_total)), clip="off")
#

@@

p2  = ggplot(data2, aes(Year, Counts)) +
geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
theme_light() +
theme(plot.margin=unit(c(0,0,0,0), "mm"),
  axis.title.y = element_text(face = "bold", size = 7, vjust = -2),
  axis.title.x = element_text(face = "bold", size = 8),
  axis.text.y = element_text(size = 6),
  axis.text.x = element_text(size = 6,  angle = 90), 
  strip.text.x = element_text(size=12),
   axis.ticks.x = element_line(color="black"),
  strip.background = element_rect(
  color="black", fill=NA, size=0.5, linetype="solid")) + 
ylab("Number of studies") +
coord_cartesian(expand = F, clip = "off") + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
panel.border = element_blank(),
axis.line = element_line(color = 'black')) +
scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
ggtitle("(b)") +
theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -5))

main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
#Select main orders
data1 = data %>% 
select(Pollinator_order, Pollinator_accepted_name) %>% 
filter(Pollinator_order %in% main_orders) %>% 
group_by(Pollinator_order) %>% 
summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.),
          Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))

# Combine Interactions and Species into a single column
data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")

#Set order of levels for plotting
data_long$Pollinator_order = factor(data_long$Pollinator_order, levels = c("Coleoptera", "Lepidoptera", "Diptera", "Hymenoptera"))

#Plot
p3 = ggplot(data_long, aes(x = Variable, y = Value, fill = Pollinator_order)) + 
geom_bar(stat = "identity", width = 0.7) +
theme_classic() +
scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator order") +
xlab(NULL) +
ylab("Percentage") + 
scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
theme(axis.text = element_text(size=6), 
     axis.title.y = element_text(face = "bold", size = 7, vjust = -2),  
     axis.title.x = element_text(face="bold", size = 7),
     legend.position = "top",
     legend.title = element_text(size = 5, face = "bold"),
     legend.text = element_text(size = 5),
     legend.key.size = unit(3, "mm"),
     legend.margin=margin(0,0,0,0),
     legend.box.margin=margin(-8,-5,-5,-5)) +
guides(fill=guide_legend(nrow=2,byrow=TRUE))+
ggtitle("(c)") +
theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6))

#plot2 = p3 / p1 / p2 
plot2 = p2 / p3

#Plot everything
map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.485, 0.8))
